Paper notes

### 2019-09-08 paper_notes.Rmd
update 03_09_2019
ok 08_09

$P_j=\frac{1}{1 + \exp(-\lambda\times \Delta_{(U_j,U_i)})}$

@@

@@

$P_i=\frac{1}{1 + \exp((\beta_ee_j + \beta_vv_j) - (\beta_ee_i + \beta_vv_i))}$

$P_j=\frac{1}{1 + \exp((-\beta_e(ej-ei)) + (-\beta_v(vj-vi)))}$

@@

The condition of random terms as independent and identically distributed

@@

The implementation of the logit model here in the agent-based model fits this condition, since the probability to

Since the logit model as formulated in the agent-based model uses the difference in the utility between the two options,

(need of the exponential function), in the conditional logit through the Gumbel distribution,  in the

have in our model where th

should be equivalent in representing a random utility model and what we need. They both are probabilities ranging from 0 to 1, the characteristic of the option regulates the probability to be selected, i.e. the utility derived from the alternative neighborhood.

This link synthetizes: [logit_binary_choice, point C](https://en.wikipedia.org/wiki/Discrete_choice#A._Logit_with_attributes_of_the_person_but_no_attributes_of_the_alternatives)

Here is the logistic function implemented in the model for the choice between current location and alternative location:

which uses the conditioanl logit to calculate the probability to choose $\textit{j}$ ($P_j=\frac{\exp(U_{(xjyj)})}{\sum_{l}exp((U_{xl,yl}))}$). As I wrote and as I understand so far, for two options the calculation of probability according to the multinomial choice formulation and the logistic function for binary choice can be considered equivalent as they both will assure the $i.i.d$ distribution of the random term, which is the priority of random utility models.\\

$k$ times ethnic weight $ew$ and $k$ times $vw$ substitute the parameters in random  utility for how important the ethnic composition and value composition are in the logistic function ($\lambda$ in $P_j=\frac{1}{1 + \exp(-\lambda\times \Delta_{U_e})+(-\lambda \times \Delta_{U_v}))}$ and $\beta_xe_j$ $\beta_yv_j$ in  $\textit{j}$ ($P_j=\frac{\exp(U_{(xjyj)})}{\sum_{l}exp((U_{xl,yl}))}$. My idea is that ethnic weight and value weight range from 0 to 1, whose value is too low causing random choices, by multiplying by a constant the ratio in the distribution point stay the same.\par
Below I report some graphs where I implement the logistic function as above tuning $k \cdot ew$ and $k \cdot vw$, checking if it would reproduce what expected from random utility models, I think it does. For parameters equal to 0, the probability to choose the alternative is $50%$ in all conditions. The best option is taken when parameters are increased, always $50%$ when the actual difference is 0. I tested some conditions considering the 2 types of utility (some would be not possible in the simulation if we keep $vw=1-ew$, but here was to see how it worked).\par
On the x-axis is ethnic utility difference, on the y-axis the value utility difference. Each ranges from -1 (alternative = 0, current = 1), to 1 (alternative = 1, current = 0), with 0 = no difference between the options. I made up them out of possible conditions. For each condition (21x21= 441), the probability function is calculated using the logisticdunction above. Heatmaps in fig. \ref{fig:heatmaps}

@@

e_w <- 0.2    # parameters to manipulate
i_v <- 0.2   # Symmetric = S:1,M:0/Threshold = S:0
S_v  <- 0

>>>>

$P_j=\frac{1}{1 + \exp(-\beta\times \Delta_{(U_j,U_i)})}$

Here is the logistic function implemented in the model for the choice between current location and alternative location and the two dimensions ethnic and value utility used:

$k$ times ethnic weight $ew$ and $k$ times $vw$ substitute the parameters in random  utility for how important the ethnic composition and value composition are in the logistic function. My idea is that ethnic weight and value weight range from 0 to 1, whose value is too low causing random choices, by multiplying by a constant the ratio in the distribution point stay the same.\par
Below I report some graphs where I implement the logistic function as above tuning $k \cdot ew$ and $k \cdot vw$, checking if it would reproduce what expected from random utility models, I think it does. For parameters equal to 0, the probability to choose the alternative is $50%$ in all conditions. The best option is taken when parameters are increased, always $50%$ when the actual difference is 0. I tested some conditions considering the 2 types of utility (some would be not possible in the simulation if we keep $vw=1-ew$, but here was to see how it worked).\par
On the x-axis is ethnic utility difference, on the y-axis the value utility difference. Each ranges from -1 (alternative = 0, current = 1), to 1 (alternative = 1, current = 0), with 0 = no difference between the options. I made up them out of possible conditions. For each condition (21x21= 441), the probability function is calculated using the logisticdunction above. Heatmaps in fig. \ref{fig:heatmaps}

@@

With utility calculated as above $\beta_ee_j + \beta_vv_j$ and $exp((U_k))$ the exponential function of the sum of all set of options

@@

$P_i=\frac{1}{1 + \exp((\beta_eUe_j + \beta_vUv_j) - (\beta_eUe_i + \beta_vUv_i))}$

From here, in terms of two options, the probability of an agent to relocate from current node \textit{i} to alternative \textit{j}  becomes:

$P_j=\frac{1}{1 + \exp((-\beta_e{(Ue_j-Ue_i)}) + (-\beta_v{(Uv_j-Uv_i)})}$

I applied it to the beta distribution of ethnic weights and value weights

@@

@@

@@

e_w <- 0.2    # parameters to manipulate
i_v <- 0.2   # Symmetric = S:1,M:0/Threshold = S:0
S_v  <- 1
